---
title: "nhanes"
author: "Hang Lei"
date: "2025-04-19"
output: html_document
---

load the package and inspect data 
```{r load libraries, message = FALSE}
library(mice)
library(haven)
library(dplyr)
library(survival)
library(ggplot2)
library(survival)
library(survminer)

```

Read the data from STATA to R. Change K6 score into categorical 
```{r}
analytic_data <- read_dta("analytic_data.dta")

analytic_data <- subset(analytic_data, analytic_sample == 1)

View(analytic_data) 

write_dta(analytic_data, "analytic_data_beforeMI.dta")



analytic_data$k6 <- ifelse(analytic_data$k6 == 0, "Non-case",
                    ifelse(analytic_data$k6 == 1, "Mild/Moderate",
                           ifelse(analytic_data$k6 == 2, "Serious", NA)))
analytic_data$k6 <- factor(analytic_data$k6, 
                           levels = c("Non-case", "Mild/Moderate", "Serious"))

```

Formatting the variables/cleaning
```{r}
DNORS <- analytic_data %>%
  select(
    returned, SRH, k6, home_damage, stuff_damage, phq9_score, ptsd, 
    homeowner, black, female, religious, edu, pre_physical_severity_cat,
    pre_mental_severity_cat, pcl_total, pss_4, fin_diff, linc, m2_impr, 
    b5, agecat, selfemployed, selfemployed_preK,months_returned
  ) %>%
  mutate(
    returned = as_factor(returned),
    SRH = as_factor(SRH),
    k6 = as_factor(k6),
    home_damage = as_factor(home_damage),
    stuff_damage = as_factor(stuff_damage),         
    phq9_score = as.numeric(phq9_score),
    ptsd = as_factor(ptsd), 
    homeowner = as_factor(homeowner),
    black = as_factor(black),                         
    female = as_factor(female),
    religious = as_factor(religious),
    edu = as_factor(edu),
    pre_physical_severity_cat = as_factor(pre_physical_severity_cat),
    pre_mental_severity_cat = as_factor(pre_mental_severity_cat),
    pcl_total = as.numeric(pcl_total),
    pss_4 = as.numeric(pss_4),
    fin_diff = as.numeric(fin_diff),
    linc = as.numeric(linc),
    m2_impr = as.numeric(m2_impr),
    b5 = as.numeric(b5),
    agecat = as_factor(agecat),
    selfemployed = as_factor(selfemployed),
    selfemployed_preK = as_factor(selfemployed_preK),
    months_returned = as.numeric(months_returned)
  )

analytic_data$pcl_total
```

Multiple imputation
```{r}
# Run Multiple Imputation (mice)
init <- mice(DNORS, maxit = 0, printFlag = FALSE)


set.seed(123)
imputed_data <- mice(DNORS,
                     m = 5,         
                     maxit = 50,    
                     method = init$method,
                     predictorMatrix = init$predictorMatrix,
                     seed = 123)


# check that imputed_data is valid mids object (successful run of imputed_data)

class(imputed_data)

comp1 <- complete(imputed_data, 1)
length(comp1)

nrow(complete(imputed_data, 2))  # Check the number of rows in the first imputed dataset

```

logistic regression model 
```{r}

# 1. Define your model using with()
model_fit <- with(data = imputed_data,
                  expr = glm(returned ~ SRH + k6 + home_damage + stuff_damage + 
                               phq9_score + ptsd + homeowner + black + female + religious + 
                               edu + pre_physical_severity_cat + pre_mental_severity_cat +
                               pcl_total + pss_4 + fin_diff + linc + m2_impr + b5 + 
                               agecat + selfemployed + selfemployed_preK, 
                               family = binomial))


# model with just physical/mental pre-existing condition 
model_fit1 <- with(data = imputed_data,
                  expr = glm(returned ~ k6 + home_damage + stuff_damage + 
                               phq9_score + ptsd + homeowner + black + female + religious + 
                               edu + pre_physical_severity_cat + pre_mental_severity_cat +
                               pcl_total + pss_4 + fin_diff + linc + m2_impr + b5 + 
                               agecat + selfemployed + selfemployed_preK, 
                               family = binomial))


# model with just self rated health 
model_fit <- with(data = imputed_data,
                  expr = glm(returned ~ SRH + k6 + home_damage + stuff_damage + 
                               phq9_score + ptsd + homeowner + black + female + religious + 
                               edu + pcl_total + pss_4 + fin_diff + linc + m2_impr + b5 + 
                               agecat + selfemployed + selfemployed_preK, 
                               family = binomial))



# 2. Pool the results across imputations
pooled_results <- pool(model_fit)

# 3. View summary
summary(pooled_results)


# Get the summary table from the pooled model
pooled_summary <- summary(pooled_results)

# Exponentiate the log-odds estimates to get odds ratios and 95% CIs
pooled_summary$OR <- exp(pooled_summary$estimate)
pooled_summary$Lower <- exp(pooled_summary$estimate - 1.96 * pooled_summary$std.error)
pooled_summary$Upper <- exp(pooled_summary$estimate + 1.96 * pooled_summary$std.error)

pooled_summary

#SRH
#home_damage/ lot of damage
#homeowner
#education Graduate degree
#self_employed


#SRH - 
#home_damage - 
#pre_katrina_home_ownership 
#race

```


Cox regression Weighted Full model(final model)
```{r}

# Cox Regression with Weights

# 1. Convertimputed_data to imputationList（for use with mitools）
library(mitools)
imputed_list <- imputationList(lapply(1:5, function(i) complete(imputed_data, i)))

# 2. Add weights to each dataset (make sure the weight column exists in analytic_data)
for (i in 1:5) {
  imputed_list$imputations[[i]]$adlt_wgt_s <- analytic_data$adlt_wgt_s
}

# 3. Create survey design object
library(survey)
design_list <- svydesign(ids = ~1, weights = ~adlt_wgt_s, data = imputed_list)

# 4. Weighted Cox model (full model version)
cox_weighted_model <- with(design_list, 
                           svycoxph(Surv(months_returned, as.numeric(returned)) ~ 
                                    SRH + k6 + home_damage + stuff_damage + 
                                    phq9_score + ptsd + homeowner + black + female + religious + 
                                    edu + pre_physical_severity_cat + pre_mental_severity_cat +
                                    pcl_total + pss_4 + fin_diff + linc + m2_impr + b5 + 
                                    agecat + selfemployed + selfemployed_preK))

# 5. Combine the results
pooled_weighted_cox <- MIcombine(cox_weighted_model)

# 6. Organize the output (Add HR and 95% CI)
summary_weighted <- summary(pooled_weighted_cox)

print(colnames(summary_weighted))


cox_formatted_weighted <- summary_weighted %>%
  mutate(
    `Hazard Ratio` = exp(results),
    `95% CI Lower` = exp(`(lower`),
    `95% CI Upper` = exp(`upper)`),
    `z` = results / se,
    `p-value` = round(2 * (1 - pnorm(abs(results / se))), 4)
  ) %>%
  select(`Hazard Ratio`, `95% CI Lower`, `95% CI Upper`, `p-value`) %>%
  tibble::rownames_to_column("Variable")

print(cox_formatted_weighted)


```

Cox regression weighted model without pre-existing condition (to see how it might confound SRH)

```{r}
# Cox Regression with Weights

# 1. Convertimputed_data to imputationList（for use with mitools）
library(mitools)
imputed_list <- imputationList(lapply(1:5, function(i) complete(imputed_data, i)))

# 2. Add weights to each dataset (make sure the weight column exists in analytic_data)
for (i in 1:5) {
  imputed_list$imputations[[i]]$adlt_wgt_s <- analytic_data$adlt_wgt_s
}

# 3. Create survey design object
library(survey)
design_list <- svydesign(ids = ~1, weights = ~adlt_wgt_s, data = imputed_list)

# 4. Weighted Cox model (full model version)
cox_weighted_model_without_pre <- with(design_list, 
                           svycoxph(Surv(months_returned, as.numeric(returned)) ~ 
                                    SRH + k6 + home_damage + stuff_damage + 
                                    phq9_score + ptsd + homeowner + black + female + religious + 
                                    edu  +
                                    pcl_total + pss_4 + fin_diff + linc + m2_impr + b5 + 
                                    agecat + selfemployed + selfemployed_preK))


# 5. Combine the results
pooled_weighted_cox_without_pre<- MIcombine(cox_weighted_model_without_pre)


# 6. Organize the output (Add HR and 95% CI)
summary_weighted_without_pre <- summary(pooled_weighted_cox_without_pre)

print(colnames(summary_weighted_without_pre))


cox_formatted_weighted_without_pre <- summary_weighted_without_pre %>%
  mutate(
    `Hazard Ratio` = exp(results),
    `95% CI Lower` = exp(`(lower`),
    `95% CI Upper` = exp(`upper)`),
    `z` = results / se,
    `p-value` = round(2 * (1 - pnorm(abs(results / se))), 4)
  ) %>%
  select(`Hazard Ratio`, `95% CI Lower`, `95% CI Upper`, `p-value`) %>%
  tibble::rownames_to_column("Variable")

print(cox_formatted_weighted_without_pre)


```



plot visualization 

Pre-existing physical condition
```{r}


# Step 1: Recode the variable (if you haven't done it yet)
analytic_data <- imputed_data %>%
  mutate(pre_physical_severity_cat = factor(pre_physical_severity_cat,
                                            levels = c(0, 1, 2),
                                            labels = c("None", "Mild (1–2)", "Moderate/Severe (3+)")))

# Step 2: Plot
ggplot(analytic_data, aes(x = pre_physical_severity_cat)) +
  geom_bar(fill = "red", color = "black") +
  labs(
    title = "Distribution of Physical Pre-existing Condition Severity",
    x = "Physical Severity Category",
    y = "Count"
  ) +
  theme_minimal()

```

Pre-existing mental condition
```{r}
# Step 1: Recode the variable (if you haven't done it yet)
analytic_data <- analytic_data %>%
  mutate(pre_mental_severity_cat = factor(pre_mental_severity_cat,
                                            levels = c(0, 1),
                                            labels = c("No", "Yes")))

# Step 2: Plot
ggplot(analytic_data, aes(x = pre_mental_severity_cat)) +
  geom_bar(fill = "blue", color = "black") +
  labs(
    title = "Distribution of  Pre-existing mental Condition Severity",
    x = "Mental Severity Category",
    y = "Count"
  ) +
  theme_minimal()

```

*Survival probability plots over time(up to 60 months)*
```{r}
surv_obj <- Surv(time = analytic_data$months_returned, event = as.numeric(analytic_data$returned))

# Fit Kaplan-Meier curve
km_fit <- survfit(surv_obj ~ 1)

# Plot survival curve (1 - return probability)
ggsurvplot(km_fit,
           data = analytic_data,
           xlim = c(0, 60),
           conf.int = TRUE,
           xlab = "Months Since Hurricane Katrina",
           ylab = "Proportion Not Yet Returned",
           title = "Time to Return Over 60 Months",
           risk.table = TRUE)

```

*Time to return status by physical condition severity* 
```{r}
fit_km <- survfit(Surv(months_returned, as.numeric(returned)) ~ pre_physical_severity_cat, data = analytic_data)

ggsurvplot(fit_km, data = analytic_data, 
           pval = TRUE, 
           conf.int = TRUE,
           title = "Time to Return by Physical Condition Severity",
           xlab = "Months Since Katrina", ylab = "Proportion Not Returned")

```

There is no clear visual difference in return timing by physical health severity. There was no significant difference in time to return across levels of pre-existing physical condition severity.(Log-rank p = 0.61)Physical condition severity doesn't appear to strongly affect return probability 
